---
title: "Proposal title"
subtitle: "Proposal"
author: 
  - name: "Vera Jackson"
    affiliations:
      - name: "School of Information, University of Arizona"
description: "Project description"
format:
  html:
    code-tools: true
    code-overflow: wrap
    code-line-numbers: true
    embed-resources: true
editor: visual
code-annotations: hover
execute:
  warning: false
---

```{r}
#| label: load-pkgs
#| message: false
library(tidyverse)
```

## Dataset

```{r}
#| label: load-dataset
#| message: false
```

The "Great American Coffee Taste Test" dataset is compiled of survey results that were filled out by participants of a taste test, hosted by James Hoffman, a world renowned barista, and coffee company Cometeer. Cometeer had sent 4 different types of coffee without labels to over 4,000 participants. The survey includes questions about coffee drinking habits, coffee preferences, individual taste test results for each of the 4 provides coffees, and individual demographics. 

I chose this data set because I was curious about the demographics and preferences of coffee drinkers. When I realized that the survey participants were those who were already customers of Cometeer, I was further interested - what are the demographics and preferences of these specific customers, and what insights does the dataset show that the company could use for marketing purposes. 

(proposal not complete at this moment) 

Make sure to load the data and use inline code for some of this information.

## Questions

The two questions you want to answer.

## Analysis plan

-   A plan for answering each of the questions including the variables involved, variables to be created (if any), external data to be merged in (if any).
