---
title: "Proposal title"
subtitle: "Proposal"
author: 
  - name: "Vera Jackson"
    affiliations:
      - name: "School of Information, University of Arizona"
description: "Project description"
format:
  html:
    code-tools: true
    code-overflow: wrap
    code-line-numbers: true
    embed-resources: true
editor: visual
code-annotations: hover
execute:
  warning: false
echo: true
---

```{r}
#| label: load-pkgs
#| message: false
library(tidyverse)
```

## Dataset

```{r}
#| label: load-dataset
#| message: false
```

The "Great American Coffee Taste Test" data set is compiled of survey results that were filled out by participants of a taste test, hosted in part by a coffee company Cometeer. Cometeer had sent 4 different types of coffee without labels to over 4,000 participants. The survey includes questions about coffee drinking habits, coffee preferences, individual taste test results for each of the 4 provides coffees, and individual demographics.

I chose this data set because I was curious about the demographics and preferences of coffee drinkers, both of customers of Cometeer and as a representation of the general North American population.

```{r}
# loading data

coffee_survey <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2024/2024-05-14/coffee_survey.csv')

head(coffee_survey)
```

## Questions

The two questions I want to answer:

1.  How do coffee drinking habits correlate with working and lifestyle situations?
2.  What combinations of coffee characteristics are the most popular?

## Analysis plan

A plan for answering each of the questions including the variables involved, variables to be created (if any), external data to be merged in (if any).

1.  Variables: cups, employment_status, number_children, wfh
    1.  I plan to create a grouped bar plot, of the count of amount of cups drunken per day ("cups"), which will be grouped by position for work status ("employment_status") and then color will indicate work from home status ("wfh"). I also want to find a way to incorporate number of children ("number_children"), as raising children is a form of labor and should be considered in lifestyle situations.
2.  Variables: strength, roast_level, caffeine
    1.  To make it easier to visualize, I will mutate the "strength", "roast_level", and "caffeine" variables from categorical to numeric variables, assigning an appropriate number scale (for example, 1-5 for "Weak" to "Very strong" strength). I then intend to create a contour plot of the three variables, visualizing what combination of the three is the most popular coffee consumed. This could be helpful to know for Cometeer, which coffee could be the most profitable to produce.
