---
title: "A Coffee Drinker's Lifestyle"
subtitle: "Proposal"
author: 
  - name: "Vera Jackson"
    affiliations:
      - name: "School of Information, University of Arizona"
description: "A data-driven analysis on how much coffee are people drinking, and what lifestyle conditions correlate."
format:
  html:
    code-tools: true
    code-overflow: wrap
    code-line-numbers: true
    embed-resources: true
editor: visual
code-annotations: hover
execute:
  warning: false
echo: true
---

```{r}
#| label: load-pkgs
#| message: false

if(!require("pacman"))
  install.packages("pacman")

pacman::p_load(tidyverse, glue, scales, here, ggthemes, janitor, ggplot2, readr)
```

## Data set

```{r}
#| label: load-dataset
#| message: false

# loading data

coffee_survey <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2024/2024-05-14/coffee_survey.csv')

coffee_survey

head(coffee_survey)

```

The "Great American Coffee Taste Test" data set is compiled of survey results that were filled out by participants of a taste test, hosted in part by a coffee company Cometeer. Cometeer had sent 4 different types of coffee without labels to over 4,000 participants. The survey includes questions about coffee drinking habits, coffee preferences, individual taste test results for each of the 4 provides coffees, and individual demographics.

I chose this data set because I was curious about the demographics and preferences of coffee drinkers, both of customers of Cometeer and as a representation of the general North American population. My first question will look at coffee drinking habits (how many cups of coffee are drunken a day) and how they correlate with working and lifestyle situations (whether they work or not, if they work from home or not, and how many children they have). It could be interesting to further analyze that data with why they drink coffee (for the caffeine, or is it for the flavor?) or if that coffee is caffeinated or not, but I want to be careful of not introducing too many variables. Overall, my goal for the first question is to highlight a trend between coffee drinking and lifestyle that could be reflective of the general American population. As for my second question, I am taking a more "marketing" approach, as I will try to determine which type of coffee (in regards to strength, roast, and caffeine-level) is most popular among Cometeer customers, as this information could assist them with product development.

## Questions

The question I want to answer:

1.  How do coffee drinking habits correlate with working and lifestyle situations?

## Analysis plan

A plan for answering each of the questions including the variables involved, variables to be created (if any), external data to be merged in (if any).

1.  Variables: cups, employment_status, number_children, wfh, age
    1.  I plan to create a grouped bar plot, of the count of amount of cups drunken per day ("cups"), which will be grouped by position for work status ("employment_status") and then color will indicate work from home status ("wfh"). I will also incorporate number of children ("number_children") into this grouped bar plot, perhaps by faceting.
    2.  As mentioned, I want to find a way to incorporate number of children ("number_children"), as raising children is a form of labor and should be considered in lifestyle situations, but the number of children may not accurately reflect how many are dependents. To address this, I will create a scatter plot, with number of children against number of cups per day, coloring by age. Incorporating "age" may further explain any trends (a 30 year old with children is more likely to have them in their home versus a 60 year old with children, although this may be an inappropriate generalization to make and I will play around with the data.) (addressing issue #1)

## Weekly Plan of Attack

| Task Name                              | Status   | Due     | Priority | Summary                                                                                     |
|----------|----------|----------|----------|-------------------------------|
| Peer Proposal                          | Complete | May 29  | High     | Provided peer reviews                                                                       |
| Final Proposal                         | Complete | June 3  | High     |                                                                                             |
| Data wrangling                         | Complete | June 7  | High     | Focusing on Question 1                                                                      |
| Attempt plots                          | Complete | June 7  | High     |                                                                                             |
| Finalize plots                         | Complete | June 14 | High     |                                                                                             |
| Begin presentation                     | Complete | June 14 | Medium   |                                                                                             |
| Finalize presentation slides and video | Complete | June 20 | High     | By this date, I want to have completed everything, leaving the weekend for any final tweaks |
| Final Presentation Due                 | Complete | June 24 | High     |                                                                                             |
